---
title: "Reproducible Research: Peer Assessment 1"
author: "Cesar Alvarez"
date: "Sunday, August 09, 2015"
output: html_document
keep_md: true
---
  
  
## Loading and preprocessing the data
### Reads activity.zip from current working directory
```{r, echo=TRUE}
actData <- read.csv(unz("activity.zip", "activity.csv"))
```

## What is mean total number of steps taken per day?  
```{r, echo=TRUE}
# Get total steps per day, ignore the missing values in the dataset
stepsPerDay <- aggregate(steps ~ date, data = actData, sum, na.rm=TRUE)

# Histogram of the total number of steps taken each day  
hist(aggregate(steps ~ date, data = stepsPerDay, sum, na.rm=TRUE)$steps, main="Total number of steps per day", 
     xlab="Steps")
```

### Mean total number of steps taken per day  
```{r, echo=TRUE}
mean(stepsPerDay$steps, na.rm = TRUE)
```

### Median total number of steps taken per day
```{r, echo=TRUE}
median(stepsPerDay$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

### Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=TRUE}
# Aggregate steps per interval and get average number of step
stepsPerInterval <- aggregate(steps ~ interval, data = actData, mean, na.rm=TRUE)
```

```{r, echo=TRUE}
# Line series plot type = l of the 5-minute interval and the average number of steps
plot(stepsPerInterval$interval, stepsPerInterval$steps, type='l', main="Average number of steps by Interval", 
     xlab="Interval", ylab="Average number of steps")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
stepsPerInterval[which.max(stepsPerInterval$steps), "interval"]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
